### 2017-03-30 ms.tex
Added some boldface and additional edits to text.  Last bit is to address the HST PSF

\documentclass{emulateapj}
\usepackage{amsmath} 
\usepackage{apjfonts} 
\usepackage{amssymb} 
\usepackage{xspace}
\usepackage{hyperref} 
\usepackage{natbib} 
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{epsf}
\usepackage{subfigure}
\bibliographystyle{apj_ads}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\comment}[1]{{\color{red} #1}}
\newcommand{\wording}[1]{{\it\color{purple} #1}}
\newcommand{\todo}[1]{{\bf\color{blue} #1}}
\newcommand{\edit}[1]{{\bf #1}}
%%%%%%%%%% User defined symbol %%%%%%
\def\gsim{\gtrsim}
\def\lsim{\lesssim}    
\def\Msun{M_\odot}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document} 
\shorttitle{Automated Lensing Learner}
\shortauthors{Avestruz, et al.}
\submitted{The Astrophysical Journal} 
\slugcomment{The Astrophysical Journal, submitted}

\title{Automated Lensing Learner - I: An automated Strong Lensing
  Identification Pipeline}

\author{Camille Avestruz\altaffilmark{1-3,$\dagger$}\thanks{E-mail:
    avestruz@uchicago.edu}, Nan Li\altaffilmark{2-4}, Matthew Lightman\altaffilmark{5}, Thomas E.~Collett\altaffilmark{6}, Wentao Luo\altaffilmark{7}}

\affil{
$^1${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^2${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^3${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.};\\
$^4${High Energy Physics Division, Argonne National Laboratory,
Lemont, IL 60439};\\
$^5${JPMorgan Chase, Chicago, IL 60603 U.S.A.};\\
$^6${Institute of Cosmology and Gravitation, University of Portsmouth, Burnaby Rd, Portsmouth, PO1 3FX, UK};\\
$^7${Center for Astronomy and Astrophysics, Shanghai Jiaotong University,
800 Dongchuan Road, Shanghai 200240, China};\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
  \href{mailto:avestruz@uchicago.edu}{avestruz@uchicago.edu}\\
}

\keywords{gravitational lensing --- methods : numerical --- methods: data analysis --- methods: statistical --- galaxies: elliptical --- surveys  }

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present (1) {\em (ALL) Automated
  Lensing Learner}, a strong lensing identification pipeline that will
be publicly released as open source software, and (2) a publicly
available mock LSST dataset with strong galaxy-galaxy lenses.  In this
first application of the pipeline, we employ a fast feature extraction
method, Histogram of Oriented Gradients (HOG), to capture edge
patterns that are characteristic of strong gravitational arcs in
galaxy-galaxy \edit{strong} lensing.  We use logistic regression to train a
supervised classifier model on the HOG of HST- and LSST-like images.
We use the area under the curve (AUC) of a Receiver Operating
Characteristic (ROC) curve to assess model performance; $AUC=1.0$ is an
ideal classifier, and $AUC=0.5$ is no better than randomly guessing.
Our best performing models on a training
set of 10,000 lens containing images and 10,000 non-lens containing
images exhibit an AUC of 0.975 for an HST-like sample.  However, for
\edit{one exposure} of LSST, our model only reaches an AUC of 0.625.  For
10-year mock LSST observations, the AUC improves to 0.809.  Model
performance appears to continually improve with the size of the
training set.  Models trained on fewer images perform better in
absence of the lens galaxy, the removal of which can decrease the
false positive rate due to the blending of light between source images
and the lens galaxy.  However, with larger training data sets,
information from the lens galaxy actually improves model performance.
Our results demonstrate an efficient and effective method for automatically
identifying strong lenses that captures much of the complexity of the
arc finding problem. The linear classifier both runs on a
personal laptop and can easily scale to large data sets on a computing
cluster, all while using existing open source tools.
\end{abstract}
%-------------------------------------------------%

%-------------------------------------------------%
\section{Introduction}
%-------------------------------------------------%

% Gravitational lensing as a geometric test of cosmology.  Also
% probes underlying matter distribution in galaxy clusters, sensitive
% to things like recent accretion and environment.
Gravitational lensing occurs when intermediate fluctuations in the
matter density field deflect light from background sources
\citep[see][for a review]{kneibandnatarajan_11}.  Strong gravitational
lensing can manifest as visible giant arcs magnifying high redshift
galaxies \citep{lyndsandpetrosian_86,gladders_etal03}, multiply imaged
quasars \citep{walsh_etal79}, multiply imaged galaxies
\citep{sharon_etal05}, and arclets \citep{bezecourt_etal98}.  Lensing
signatures probe the underlying dark matter distribution of the lens
\citep{warrenanddye_03}, and high redshift galaxy formation
\citep{allam_etal07}. Strong lenses also provide a geometric test of
cosmology via comparison of predicted arc abundances with observed
abundances \citep{kochanek_96,chae_03,linder_04}, via time-delay
between signals from multiply imaged quasars \edit{
\citep{xli_etal12,suyu_etal14,suyu_etal16,bonvin_etal17}, and via
distance ratios in lenses with sources at multiple redshifts
\citep{jullo_etal10,collett_etal12,collettandauger_14}.}

% Application of constraining mass profiles - large numbers needed.
The application of strong gravitational lensing to constrain the mass
distribution of strong lenses, such as early-type galaxies (ETGs),
necessitates large samples of galaxy-galaxy \edit{strong} lensing systems.
\citet{miraldaescudeandlehar_92} first suggested that massive
ellipticals would likely be frequent strong lensing sources in optical
surveys.  These systems contain a background source galaxy that the
lens galaxy deflects into a partial or full arc shaped Einstein ring.
The strong lensing signature directly probes the underlying matter.
The identification of such systems in upcoming surveys is the first
step in constraining the mass-to-light ratio for a large number of
objects in this mass range.

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure for both large scale visual and
automated image classification emerged. By nature, the human eye is
one of the best discriminators for image classification.  Visual arc
identification has been effective through the use of citizen science
platforms.  {\em SpaceWarps} is an example of citizen science based
image classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  These platforms are quite
successful for a dataset like CFHTLS; here, 3,000 candidate images
were identified in eight months, resulting in 89 final candidates.
However, future datasets like Euclid \citep{oguriandmarshall_10} and
LSST expect to find \edit{hundreds of thousands galaxy-scale strong lenses
\citep{collett_15}.}  The volume of upcoming data challenges the
scalability of a pure citizen science approach.

% Transition to automated identification
Recent efforts have focused on the development of automated methods
with performance comparable to or better than humans. {\em SpaceWarps}
is a part of the Zooniverse Project \citep{marshall_etal16}, which
also includes {\em Galaxy Zoo}, the citizen science based image
classification of galaxy types \citep{lintott_etal08}.  Galaxy
classification is an early example in which machine
learning algorithms successfully trained models to classify astronomical
images with comparable performance to humans \citep{dieleman_etal15}.

% Early efforts
Earlier efforts on automated, or ``robot'', identification of strong
lensing systems have two distinct generalized steps.  The first
enhances and extracts characteristic features, and the second uses
some form of pattern recognition in the features to classify lens and
non-lens containing systems.  Among others, selected features might
include shape parameterization
\citep[see][e.g.]{,alard_06,kuboanddellantonio_08,xu_etal16,lee_17},
colors in multi-band imaging \citep{maturi_etal14,gavazzi_etal14},
light profiles \citep{braultandgavazzi_15}, and characteristics of
potential lens galaxies \citep{marshall_etal09}.  Pattern recognition
often incorporates cutoffs in selected parameter space distributions
\citep{lenzen_etal04,joseph_etal14,paraficz_etal16}.  A number of
these have been publicly distributed, with specific end applications.
For example, {\em ArcFinder} is one such code that finds arcs in
groups or cluster scale lens \citep{seidelandbartelmann_07}, and {\em
  RingFinder} is an analogous tool in searching for multiply imaged
quasars \citep{gavazzi_etal14}.  Codes like these have been
complementary to human identification \citep{more_etal16}.

% Classical machine learning methods
Pattern recognition methods have transitioned to using ``machine
learning'' algorithms in place of manual cutoffs.  With machine
learning, we can train a model to separate a dataset into a known set
of classes, such as ``lensing systems'' and ``non-lensing
systems''. However, many classic machine learning algorithms do not
work well in image space, i.e. directly using the raw pixel values of
the image, but first require a process of {\em feature engineering}.
{\em Feature engineering} uses domain expertise to extract variables
that are more directly related to the classification task at hand.  An
example would be the application of an edge detector.  The optimal
weights and cutoffs for these derived variables that are used to
determine the class label of an image are found automatically by the
algorithm.

% Recent works, neural nets, and deep learning
Some more recent works have made use of a subset of machine
learning algorithms called neural networks to classify images,
either from derived image
parameters or directly in image space.  In \citet{estrada_etal07},
authors used derived shape parameters to train neural nets to identify
arc candidates.  \citet{agnello_etal15} used neural networks trained
on data from multiband magnitudes, and \citet{bom_etal16} used
extracted morphological parameters.

% Convolutional and residual
The most advanced use of neural networks for strong lensing
classification operate directly in image space.
\citep{petrillo_etal17} used mock Kilo Degree Survey data to train
convolutional neural networks, with a training set size of six million
images. \citet{lanusse_etal17} used state-of-the-art deep residual
neural networks to also work directly in the image space with minimal
image pre-processing.  A major strength of the \citet{lanusse_etal17}
implementation is that in comparison to deep convolutional neural
networks, deep residual neural networks have been found to be easier
to train and perform better (Metcalf et al., in prep).

%  Relevance of our work
While there has been a recent surge in the use of deep neural networks
applied to image classification problems in astronomy, it is not
always easy to scale these techniques to large data sets, nor are the
necessary computational resources and hardware, such as graphical
processing units (GPUs), easily accessible to the entire scientific
community.  We present a supervised classification pipeline built with
open source tools.  We test the pipeline on mock HST and LSST data,
which will also be made publicly available.  Our goal is to create a
centralized open source pipeline that is general enough for the user
to select or add appropriate image processing techniques to augment
the features of the lensed image, train and test data with different
machine learning techniques, and to select the sequence of methods
that provides optimal model performance.  Results are reproducible on
personal computers, as both the pipeline and the data will be publicly
distributed.

% Summary of paper 
This first paper serves as both a presentation of the mock data set
and as a thorough case study for our pipeline with one feature
extraction and classical machine learning method, respectively, the
{\em histogram of oriented gradients} (HOG) and {\em logistic
  regression} (LR).  HOG is a fast feature extraction procedure that
quantifies edges in images, commonly used to identify humans in
security software.  LR is a linear classifier, making its scalability
relatively straightforward with existing open source tools.  This
framework is easily extendable to other feature extraction methods
and machine learning algorithms for supervised classification.

% Summary of results
We show results of the pipeline on mock galaxy-galaxy lens systems
observed by HST and LSST as respective examples of classifier
performance on optical space- and ground-based observations.  We train
and test our pipeline on subsamples from 10,000 mock observed strong
lensing systems and 10,000 non-strong lensing systems, each centered
on a potential lens galaxy.

Our paper is organized as follows. In Section~\ref{sec:methods} we
briefly describe the methods to generate the mock HST and LSST data
and our overall image processing and classification pipeline. We
present our results in Section~\ref{sec:results}, and our summary and
discussions in Section~\ref{sec:conclusions}.

%-------------------------------------------------%
\section{Methodology}
\label{sec:methods}
%-------------------------------------------------%
\subsection{Mock Images}
%-------------------------------------------------%
We use mock {\em Hubble Space
  Telescope}\footnote{https://www.spacetelescope.org/} (HST) and {\em
  Large Synoptic Sky Telescope}\footnote{https://www.lsst.org} (LSST)
images generated with {\em PICS} (Pipeline for Images of Cosmological
Strong lensing) \citep{li_etal16} and {\em LensPop} \citep{collett_15}
for training and testing.

There are 10,000 lens containing mock observations and 10,000
non-lens containing mock observations for running our parameter search. 
We keep a hold-out set of 1,000 lens containing mock observations
and 1,000 non-lens containing mock observations on which we evaluate
the final trained model.

Mock observations of lensing systems include the lens galaxy, lensed
images of the source galaxy, and galaxies along the line of sight.
Mock observations of non-lensing systems include all but the images of
a lensed source galaxy.  We convolve each of the $2\times 10,000$
train/test images and $2\times 1,000$ hold-out images to produce three
separate sets of mock ``observations''.  These each have equal numbers
of lens and non-lens containing systems: (1) HST-like observations,
(2) best single exposure LSST-like observations, and (3) LSST-like
observations over the span of ten years.  We respectively label these
observations as HST, \edit{LSST-best}, and LSST10.

%-------------------------------------------------%
\subsubsection{Modeling the Mass Distribution of Lens Galaxies}
%-------------------------------------------------%

To produce simulated lensed galaxies, we first model the mass of lens
galaxies as a Singular Isothermal Ellipsoid (SIE). This model is
both analytically tractable and is consistent with models of
individual lenses and lens statistics on the length scales relevant
for strong lensing
\citep[e.g.]{koopmans_etal06,gavazzi_etal07,dye_etal08,liandchen_09}.

The normalized convergence map of the SIE model is defined as:
%-------------------------------------------------%
\begin{equation}\label{eqn:sie}
\kappa = \frac{\theta_{\rm E}}{2}\frac{1}{\sqrt{x_1^2/q+x_2^2 q}},
\end{equation}
%-------------------------------------------------%
where $\theta_{\rm E}$ is the Einstein Radius and $q$ is the axis ratio.
The Einstein radius can be calculated from the redshift of the
lens, the redshift of the source, and the velocity dispersion of the
lens galaxy as follows,
%-------------------------------------------------%
\begin{equation}
\theta_{\rm E} = 4\pi\left(\frac{\sigma_v}{c}\right)^2\frac{D_{ls}}{D_{s}}, 
\end{equation}
%-------------------------------------------------%
Here, $c$ is the speed of light, $\sigma_v$ is the velocity dispersion
of the lens galaxy, $D_{ls}$ and $D_{s}$ are respectively the angular
diameter distances from the source plane to the lens plane and from
the source plane to the observer.

To rotate the lenses with random orientation angle $\phi$, we adopt
the transformation below:
\begin{eqnarray}
\label{eqn:rotation}
\begin{bmatrix}
    x_{1}\prime      \\
    x_{2}\prime 
\end{bmatrix}
= 
\begin{bmatrix}
    \cos \phi  & -\sin \phi      \\
    \sin \phi  & ~\cos \phi      
\end{bmatrix} 
\begin{bmatrix}
    x_1      \\
    x_2      
\end{bmatrix}.
\end{eqnarray}

From Equations~\ref{eqn:sie}~-~\ref{eqn:rotation}, there are five
independent parameters from which we can derive the lensing map:
velocity dispersion $\sigma_v$, axis ratio or ellipticity $q$,
orientation angle $\phi$, redshift of the lens $z_l$, and redshift of
the source $z_s$.

We choose $\sigma_v$, $q$, and $\phi$ \edit{randomly (flat prior) from
typical ranges of observed galaxies: $\sigma_v \in [200, 320]~km/s$,
$q \in [0.5, 1.0]$, and $\phi \in [0, 360]$.  While a flat prior is
not realistic, we use this as a starting point to test our pipeline
(see Section~\ref{sec:conclusions} for more details on future
work)}. We obtain the redshift of the lens galaxy by matching the
velocity dispersion from our simulations to the catalog of elliptical
galaxies in the COSMOS survey from \citet{zahid_etal15}, resulting in
lens galaxy redshifts in the range, $z_l \in [0.2,0.7]$.

%-------------------------------------------------%
\subsubsection{Modeling Images of the Lens Galaxies}
%-------------------------------------------------%
We model the light distributions of the lens galaxies with an
elliptical Sersic profile,
%-------------------------------------------------%
\begin{equation}\label{eqn:sersic} 
I(R) = I_{\rm eff}~{\rm exp} \left\lbrace -b_{n} \left[ \left(
  \frac{R}{R_{\rm eff}}\right)^{1/n} - 1 \right ]
\right\rbrace 
\end{equation}
%-------------------------------------------------%
where, $R = \sqrt[]{x_1^2 /q+x_2^2 q }$, $R_{\rm eff}$ is the
effective radius in units of arcseconds, $I_{\rm eff}$ is the
intensity at the effective radius, $n$ is the index of the Sersic
profile, and $q$ is the ellipticity of the lens galaxy.  We perform a
similar transformation as Equation~\ref{eqn:rotation} to orient the
source galaxies and assume that the distribution of light follows that
of mass.  The ellipticity and orientation are therefore the same as in
the SIE model.

We use the COSMOS morphological catalog \citep{zahid_etal15} to match
the velocity dispersion with an assigned effective radius, effective
luminosity, and index to the light profile.  We also assume the light
center is on top of the mass center, creating noiseless images of lens
galaxies.

We construct galaxies along the line of sight by cutting light-cones
from the Hubble Ultra Deep
Field\footnote{http://www.spacetelescope.org/science/deep\_fields/}.
We stack these images with the lens galaxy image and the lensed source
galaxy, calibrating the magnitude of all three components.

%-------------------------------------------------%
\subsubsection{Modeling Images of the Source Galaxies}
%-------------------------------------------------%
The background source galaxies come from a set of detailed images of
low-redshift bright galaxies ($z\sim0.45$) that have been extracted
from mosaics produced by the CANDELS team
\citep{grogin_etal11,koekemoer_etal11} and selected from the CANDELS
UDS catalog \citep{galametz_etal13}.  We rescale these background
source galaxies, set the source redshift to \edit{$z_s=2$, and select
  source positions near caustics of the lensing system.  The selected
  positions are within a $2''\times2''$ square box centered on the
  lens galaxy center of mass.} We then fix the projected position of
the lens galaxy at the center of the field of view for each image.

We perform ray-tracing simulations using the modeled galaxy images and
parameters.  The lens galaxy, source galaxy, and parameters of the
image simulations (i.e., $300 \times 300~{\rm pixels}^2$ with
0.03~arcsec/pixel for an HST-like image) are inputs to the simulation
that produce ideal lensed images with the appropriate resolution.

%-------------------------------------------------%
\subsubsection{Mock Observing}
%-------------------------------------------------%

The final part is to perform the mock observation from the ideal
images produced in the previous step.  \edit{Both our HST- and
  LSST-like observations are monochromatic, respectively using the
  F606W and the g-band.} For HST-like observations, galaxies along the
line of sight already include noise and a point spread function (PSF).
We therefore stack and magnitude calibrate all components to produce
the final HST-like images, which are $300\times300~{\rm pix}^2$ with
$0.03$~arcsec per pixel.

For LSST-like observations, we use the {\em LensPop} software
\citep{collett_15}.  We resample images to match the detector pixel
scale and convolve the resampled image with a circularly symmetric
Gaussian Point Spread Function discretised at the same pixel scale. To
generate a noisy realization of the image, we assume a Poisson model
based on the sky plus signal, and an additional Gaussian read-out
noise. Parameters for these simulations follow \citet{collett_15} and
are based on the LSST observation simulator \citep{connolly_etal10}.

To account for variations in seeing and sky-brightness over the course
of the survey, we draw each simulated exposure from a stochastic
distribution of these parameters. We then consider two different
strategies to use the simulated exposures. First, we build one
single-epoch image for each field (hereafter labeled as \edit{LSST-best}) by
keeping only the best seeing exposure.  Second, we build another
``worst-case'' stacked image by degrading all individual exposures,
ten per filter per year, to match the one with the worst seeing and
co-add all exposures to a single image (hereafter labeled as LSST10).
These two sets of images will allow us to investigate the trade-off
between resolution and signal-to-noise for our automated lens search.

Figure~\ref{fig:mockimages} illustrates sample mock observations with
a strong lensing signature from each telescope.  The left-most column
corresponds to a mock HST lensing system with a highly magnified
source galaxy (top) and a less visible image of the source galaxy
(bottom). For the HST-like dataset, many arcs are visually obvious due
to the exquisite spatial resolution and quality of space based
imaging.

The middle column corresponds to the same simulated systems for
LSST10.  The right-most column corresponds to the simulated systems of
\edit{LSST-best}.  These images have resolution $45\times45~{\rm pixels}^2$ with
$0.18$~arcsec/pixel.  LSST10 images visually exhibit the improved
signal to noise ratio, recovering the arc feature, albeit at a much
lower resolution than with the HST-like image or the \edit{LSST-best} image.
The top images of LSST10 and \edit{LSST-best} show a visible lensed source
galaxy image.  The bottom images are washed out in the bottom row,
where the magnification of the source galaxy is not as large.  The
ground based noise, PSF, and limited resolution of the \edit{LSST-best} make
visual giant arc identification difficult, except in systems with the
most magnified source galaxies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{figures/tiles/compilations/telescope_compilation.pdf}
\caption{Left to right show example mock HST, LSST 10 year, and
  \edit{LSST-best} images.  The top row corresponds to a lensing system with a
  very visible arc signature, and the bottom row to a lensing system
  that is less obvious.  Example mock HST images have
  $n_\text{pix}\times n_\text{pix}=300\times300$.  Example mock LSST
  images have $n_\text{pix}\times n_\text{pix}=45\times45$.  The
  resolution and noise of a ground based telescope is noticeably
  worse. Visual identification of giant arcs in the LSST images in the
  bottom row is very difficult.}\label{fig:mockimages}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Our mock observations also include non-lens containing images.  The
procedure is similar to mock lensed images but we do not perform
ray-tracing, so these images do not have lensed source galaxies.

Furthermore, we investigate the influence of light from the lens
galaxies on the performance of our lens identification pipeline.  We
generate another set of each HST, \edit{LSST-best}, and LSST10 images without the
lens galaxy.  We respectively label these nHST, n\edit{LSST-best}, and nLSST10.

The final data set is then comprised of $6\times10,000$ lens
containing images, and $6\times 10,000$ non-lens containing images.
We also keep a hold-out set of $6\times 1,000$ lens containing images,
and $6\times 1,000$ non-lens containing images.

%-------------------------------------------------%
\subsection{Strong Arc Lensing Identification Pipeline}
%-------------------------------------------------%

To perform our analysis, we have used tools from {\em Scikit-learn}
\citep{pedregosa_etal12}.  We outline the identification pipeline in
Figure~\ref{fig:pipeline}, which is a general description of
supervised classification.  Supervised classification is a class of
machine learning where the class labels in the training
set are known.  In our case, the labels are ``lens'' and ``non-lens''
containing images.

The first step of our pipeline consists of a feature extraction stage,
where our feature vector is a {\em histogram of oriented gradients}
(HOG) \citep{dalalandtriggs_05} that quantifies edges in the image.  We
describe the method and parameter search in Section~\ref{sec:hog}.  We
then use {\em Logistic Regression} (LR), a machine learning algorithm
described in Section~\ref{sec:LR}, to train a classifier model on a
subset of our images.  LR requires a parameter search over the
regression coefficient, $C_{LogReg}$, which we explore in
Section~\ref{sec:regularization}.  We briefly comment that our initial
tests with a {\em Support Vector Machine} (SVM) using radial basis
functions as an alternative
machine learning algorithm yielded negligible performance improvement,
and significantly increased computation cost.  This indicated that the
features of lens and non-lens images are relatively well separated by
hyperplanes in feature space.  For these reasons, we do not include
SVM in our final analysis and comparisons and continue all discussions
with a linear classifier.

Both the feature extrator, HOG, and the linear classifier, LR, contain
parameters, which must be tested and optimized for peak model
performance.  We use {\em GridSearchCV} from {\em Scikit-learn} to
select cross-validated parameters, and discuss this step of our
methodology in Section~\ref{sec:gridsearch}.

The second step of our analysis is to test our trained model on an
independent subset of the images to assess the model performance.
Here, we evaluate the model on each test image, predicting a
likelihood (``score'') between 0 and 1 that image contains a lensing
system.  This ``holdout set'' is not used in any of our parameter
searches to keep our test metric independent of tuning.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:pipeline}
\begin{center}
\includegraphics[scale=0.55]{figures/supervised_classification_workflow.png}
\caption{Cartoon of pipeline \footnote{from
    http://www.nltk.org/book/ch06.html} for supervised classification.
  In our case, the labels are either lens or non-lens containing
  images.  Our input is the set of mock HST- or LSST-like
  observations.  The feature extractor is the histogram of oriented
  gradients, producing an N-dimensional feature vector that quantify
  edges in our images.  Our machine learning algorithm is logistic
  regression.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\subsubsection{Feature Extractor: Histogram Oriented Gradients}\label{sec:hog}
%-------------------------------------------------%
% HOG intro - 
Originally created for human detection in computer vision, histogram of
oriented gradients (HOG) is a feature extraction method that computes
centered horizontal and vertical gradients. HOG is relatively robust
to noise in the image, and is a fairly fast transform that describes
edges.  Details can be found in \citet{dalalandtriggs_05}, but we
describe the procedure here.  The end result of HOG is a one
dimensional histogram computed as follows.

% Division of the image  
HOG first divides the image into blocks of 50\% overlap.  Each block
contains $m\times m$ cells-per-block that each contain $n_{pix}\times
n_{pix}$ pixels-per-cell.  The computed gradient orientation is
quantized into $N_{orient}$ bins.

% Initial binning
The orientation gradient of all pixels within each cell are binned
into the quantized orientations, providing a net gradient description
within that cell.  As an example, for $N_{orient}=3$, our bins are
centered at $\theta=0, 2\pi/3, 4\pi/3$ in radians.  If a cell
only has a gradient in the
$\theta=\pi/2$ direction, it will contribute 75\% of its magnitude to the
$\theta=2\pi/3$ bin, and 25\% of its magnitude to the $\theta=0$ bin.
The bins in all cells are then concatenated to make a larger feature
vector that is $N_{orient}$\times$N_{cells}$.

% Normalization
The last step is a normalization procedure to control for illumination
effects.  Here, the sub-histograms of each cell within the same block
are normalized with respect to one another before the transformation
returns the final feature vector.  The division of the image and the
quantization of orientations are thus controlled by 3 parameters in HOG:
$N_{orient}$, cells-per-block, and pixels-per-cell.  We discuss how we
select parameters using cross-validation in
Section~\ref{sec:gridsearch}

%-------------------------------------------------%
\subsubsection{Optimized Pipeline Parameters with a Grid Search}\label{sec:gridsearch}
%-------------------------------------------------%

\begin{table}
\caption{Grid Search of Pipeline Parameters}
\begin{center}
\begin{tabular}{cccccc}
\hline \\ [-0.2cm]
N$_{orient}$ & Pixels/Cell & Cells/Block & C$_{reg}$ & Score & \\ [0.2cm]
\hline \\ [-0.2cm]
\multicolumn{6}{c}{(a) HST-like data} \\ [0.2cm]
\hline \\ [-0.2cm]
9 & (16, 16) & (4, 4) & 10. & $0.90141\pm 0.00658$ &\\ [0.2cm]
9 & (24, 24) & (3, 3) & 10. & $0.89266\pm 0.01033$ &\\ [0.2cm]
9 & (16, 16) & (1, 1) & 10. & $0.87484\pm 0.00671$ &\\ [0.2cm]
7 & (16, 16) & (4, 4) & 10. & $0.90031\pm 0.00956$ &\\ [0.2cm]
5 & (16, 16) & (4, 4) & 10. & $0.89453\pm 0.00840$ &\\ [0.2cm]
9 & (32, 32) & (4, 4) & 10. & $0.85828\pm 0.01004$ &\\ [0.2cm]
9 & (32, 32) & (4, 4) & 50. & $0.87344\pm 0.01267$ &\\ [0.2cm]
\hline \\ [-0.2cm]
\multicolumn{6}{c}{(b) LSST-like data} \\ [0.2cm]
\hline \\ [-0.2cm]
3 & (5, 5) & (4, 4) & 100. & $0.64328\pm 0.00769$ &\\ [0.2cm]
4 & (5, 5) & (4, 4) & 100. & $0.62469\pm 0.00485$ &\\ [0.2cm]
3 & (5, 5) & (2, 2) & 100. & $0.63281\pm 0.00869$ &\\ [0.2cm]
5 & (5, 5) & (4, 4) & 100. & $0.62734\pm 0.00800$ &\\ [0.2cm]
3 & (3, 3) & (4, 4) & 100. & $0.59109\pm 0.00313$ &\\ [0.2cm]
3 & (3, 3) & (2, 2) & 100. & $0.57297\pm 0.00508$ &\\ [0.2cm]
4 & (4, 4) & (3, 3) & 100. & $0.63125\pm 0.01227$ &\\ [0.2cm]
\hline
\end{tabular}
\end{center}
\label{tab:gridsearch}

\tablecomments{Panel (a) shows a subsample of the results of a grid
  search for HST across a range of HOG parameters and regression
  coefficient. Panel (b) shows a subsample of the results of a grid
  search for LSST10. Each use a data set of size $2\times8000$ for
  cross validation to get the average scores and standard deviation.
  We explore different HOG parameters in each dataset due to
  resolution and image size differences.}
\end{table}

We run a grid search across parameters that should reasonably sample
the arc edges in either the HST- or LSST-like mock observations, and
illustrate the results in Table~\ref{tab:gridsearch}.  Recall, the
HST-like images are $300\times300$ pixels per image, while the
LSST-like mock observations are $45\times45$ pixels per image.

We first estimate the size of a cell that will contain a coherent arc
feature.  To first order approximation, subdivisions of cells that are
1/100th the area of the entire image should contain coherent arc edges
that span an elongated shape within arc-containing cells.  Therefore,
we sample the pixels per cell parameter from (8, 8) to (32, 32) for
the HST-like images, and (3, 3) to (5, 5) for the LSST-like images in
our grid search.

Next, the cells per block parameter determines the normalization of
each cell with respect to the neighboring cell.  In general, this will
down-weight arc-like edges in cells that neighbor very bright cells,
such as cells that cover the central lens galaxy.  We therefore vary
the cells per block parameter between (2, 2) and (4, 4) for the
LSST-like images and between (1, 1) and (4, 4) for the HST-like
images.

The number of orientations will determine the sampling of rounded
edges.  For example, if we only have two orientations, an arc-like
feature in a cell directly north of the lensing galaxy will appear in
our HOG visualization as a strong horizontal line (e.g. see top left
in Figure~\ref{fig:hogimages}), and an arc-like feature north-east of
the lensing galaxy will appear as an L-shape.  However, contributions
from a cluster or line-of-sight galaxy in the same cell will tend to
contribute edges in all orientations of the histogram (e.g. bottom
right in Figure~\ref{fig:hogimages}).

Finally, the resolution of the overall image will also limit the
additional information that an increase in $N_\text{orient}$ will
provide.  From the grid search, the best case number of orientations
for each dataset is $N_\text{orient,HST}=9$,
$N_\text{orient,LSST10}=3$, and $N_\text{orient,\edit{LSST-best}}=5$.

% Impact on image performance
The image resolution affects the length of the HOG feature vector,
which has a monotonically increasing relationship with the time
required to train the model.  Additionally, for fixed memory
restrictions, there is a tradeoff between the length of the feature
vector and the size of the training set.  We will discuss how the
training set size affects the train time for each data set in
Section~\ref{sec:trainsetsize}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{figures/tiles/compilations/hog_telescope_compilation.pdf}
\caption{Left to right show example image transforms of mock images
  from Figure~\ref{fig:mockimages} with a visualized histogram of
  oriented gradients.  The image transform picks up edge features,
  with arc features showing up as edges across radial orientations.
  Each of the oriented gradients within a cell is color-coded by
  magnitude, and represented as a line in the direction perpendicular
  to that gradient.  The actual extracted features fill a
  one-dimensional feature vector comprised of the magnitudes of each
  of the oriented edges within the visualized
  cells.}\label{fig:hogimages}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%% \caption{Top row: Visualized HOG feature vectors of the sample
%%   lensed HST-like image from Figure~\ref{fig:mockimages}.  Left:
%%   The corresponding HOG visualization with N$_\text{orient}=5$,
%%   (32, 32) pixels per cell, and (1, 1) cells per block.  Right: The
%%   HOG visualization of the same mock image with
%%   N$_\text{orient}=9$, (16, 16) pixels per cell, and (1, 1) cells
%%   per block.  The brightest lines align with prominent edges.
%%   Note, while fewer pixels per cell will better resolve the arc
%%   feature, the ``edge'' corresponding to the arc is less prominent.
%%   Bottom row: Visualized HOG feature vectors for the LSST-like
%%   single epoch and stacked images with the best scoring HOG
%%   parameterization of N$_\text{orient}=4$, (3, 3) pixels per cell,
%%   and (3, 3) cells per block.  The single epoch image (left) does
%%   not have significant enough edge features to emerge in the HOG,
%%   but the stacked image (right) does.}  \end{center} \end{figure*}
%%   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\subsubsection{Machine Learning Algorithm: Logistic Regression}\label{sec:LR}
%-------------------------------------------------%
The problem of detecting gravitational lenses in images falls under
the general category of {\em{classification}} in machine learning. In
general, the task is to find a function that assigns data points $x$ to
one of two or more classes, denoted by the class label $y$.  This is
equivalent to specifying a decision boundary, or decision boundaries
between the classes in the space of the data points.  (Compare this to
{\em{regression}} in which the task is to find a function $y=f(x)$,
where $y$ is a continuous, rather than discrete, variable.) In our
case, we have two classes: lens and non-lens containing images, and
the data points $x$ are the HOG feature vectors extracted from the
images.  In this paper we use the {\em Logisitic Regression}
(LR) algorithm, for which the decision boundary is a hyperplane.  (The
equivalent in the regression setting would be linear regression.) In
LR we determine the optimal hyperplane by minimizing the objective
function
\begin{equation}
\label{eqn:log_reg_objective}
L(A,b) = \sum_i \log \left[1 + \exp \left(-y_i (A \cdot x_i + b) \right) \right]
\end{equation}
where $x_i$ is a data point (HOG feature vector), $y_i$ is the
known label for that data point ($1$ for a lens containing
image, $-1$ for a non-lens containing), and $A$ and $b$ are
the parameters of the hyperplane. Eq. \ref{eqn:log_reg_objective}
is to be minimized with respect to $A$ and $b$.
Other more complicated machine learning algorithms exist which
do not necessarily produce a linear decision boundary, such as
{\em Support Vector Machines} (SVM), {\em Random Forests}, and
{\em Neural Networks} \citep{hastie_09}.

The HOG feature vectors in this paper can be very high-dimensional.
When dealing with high-dimensional data, where the number of
dimensions becomes comparable to the number of data points,
overfitting can become an issue. (An example of an extreme case
of overfitting is fitting a degree $n$ polynomial to $n$
points.  The polynomial would simply wiggle so that it goes
through every point, and would have no predictive power if you
tried to interpolate or extrapolate.) In machine learning,
overfitting is avoided using different {\em{regularization}}
techniques. A common choice for logistic regression is to
add a penalty term to Equation~\ref{eqn:log_reg_objective}
\begin{equation}\label{eqn:log_reg_regularized}
L_\text{Reg}(A,b,C_\text{LogReg}) = L(A,b) + \frac{1}{2C_\text{LogReg}} \|A\|
\end{equation}
where the norm $\|A\|$ is typically taken as either the
sum of squares of the coefficients ($L_2$ norm) or the sum
of absolute values of the coefficients ($L_1$ norm).  In this
paper we use the former.

The amount of regularization is controlled by the parameter
$C_\text{LogReg}$: larger values of $C_\text{LogReg}$ correspond to
increasing model complexity.  If $C_\text{LogReg}$ is too large then
the model will overfit, and if it is too small the model will
underfit.  To determine whether a model is overfit or underfit, the
model is trained, (i.e. Eq. \ref{eqn:log_reg_regularized} is
minimized), on a subset of the data called the {\em{training set}} and
its performance (goodness of fit) is evaluated both on the
training set and a separate {\em{test set}}
that was not used in constructing the model.
Figure~\ref{fig:regularization} shows the performance of a model with
selected HOG parameters as a function of the regularization parameter,
$C_\text{LogReg}$, for both the training and the test set.  (The
performance can be measured by the accuracy, i.e. percent of images
correctly classified, or by some other metric, such as the area under
the Receiver Operating Characteristic curve described in
Section~\ref{sec:ROC}.)

When $C_\text{LogReg}$ is small, the performance of the model improves
with increasing $C_\text{LogReg}$ on both the training and test set,
meaning that $C_\text{LogReg}$ is
still in the underfitting regime. Eventually the performance on the
test set reaches a maximum and starts to decrease, even while the
performance on the training set continues to increase.  This means
that the model is no longer generalizing well and is starting to
overfit. The optimal $C_\text{LogReg}$ occurs when the performance
of the test set is at its maximum; this is the value of $C_\text{LogReg}$
that should be used in the final model.

In practice, there is something of a trade-off between accuracy and
computational resources because a larger value of $C_\text{LogReg}$
will also increase the training time, since a larger $C_\text{LogReg}$
corresponds to a less constrained parameter space being searched.
We discuss the performance and training time dependence on $C_\text{LogReg}$
in Section\ref{sec:regularization}.

%-------------------------------------------------%
\section{Results}
\label{sec:results}
%-------------------------------------------------%

%-------------------------------------------------%
\subsection{Receiver Operating Characteristic}\label{sec:ROC}
%-------------------------------------------------%

In this section, we discuss the Receiver Operating Characteristic
(ROC) curve (see Figure~\ref{fig:ROCcompilation}), which shows the
true positive rate (tpr) as a function of false positive rate (fpr)
for a given model and test set.  The true positive rate is defined as
the number of true positives divided by the total number of positives
(completeness of the positive identifications).  The false positive
rate is defined as the number of false positives divided by the total
number of negatives (impurity of the identifications).  The ROC curve
illustrates the performance of our trained model as we vary the
discrimination threshold.

The classifier model assigns a score to each test image, which is a
probability that the image is a strong lensing system.  To construct
the ROC curve, we rank the test images by probability, and calculate
the tpr and fpr for decreasing discrimination threshold.

For very high discrimination threshold, we have maximal purity, but a
low completeness (bottom left region of the ROC).  For a very low
discrimination threshold, we have maximal completeness, but minimal
purity (top right region of the ROC).  The ideal model would have an
ROC curve with data points that go from $(x, y) = (0, 0)$ to $(x, y) =
(0, 1)$ to $(x, y) = (1, 1)$.

In the context of strong lensing systems, we wish to maximize the true
positive rate so we have a representative count of the fraction of
strong lensing systems in an observed volume of the universe.  We also
want to minimize the false positive rate.  Positively identified
strong lensing systems will require expensive spectroscopic follow-up
for validation.  The steepness of the ROC curve indicates how well the
model will optimize the two, and we can characterize the steepness by
the area under the curve (AUC).  The ideal model would have an AUC of
1.  We show the ROC curves of our best performing models in each
dataset.

% Precision-recall
For comparison, we also show a Precision-Recall curve (see
Figure~\ref{fig:PRcompilation}), where precision is the number of true
positives identified as positive divided by the total number of
positively identified images (purity) and recall is the number of true
positives identified as positive divided by the total number of true
positives (completeness).  Each point in the figure is calculated with
a varying threshold for identification.  Since our sample has a class
balance of 50-50 between positives (lens) and negatives (non-lens),
the most lenient threshold for identification would yield a precision
of 0.5 at a recall of 1.0.  It is important to note that this figure
changes as the class balance changes; if we had 90\% non-lenses and
10\% lenses, the most lenient threshold would yield a precision of 1/9
at a recall of 1.  The ROC curve is a more typical metric for
supervised classification.

% Summary of ROC curves
Figure~\ref{fig:ROCcompilation} shows the ROC curves for models
trained using the entire 10,000 training sample, with best-case HOG
and regularization parameters.  The models have been evaluated on a
hold-out set of 1,000 images that were not used in the parameter
search.  We show the mock HST, \edit{LSST-best}, and LSST10 results respectively
in red, blue, and green.  Solid lines correspond to a model trained
and tested on images with the lensing galaxy.  Dashed lines correspond
to a model trained and tested on images where the lensing galaxy is
excluded from the mock observation, simulating an ideal modeling and
subtraction of the lensing galaxy, which has been one proposed method
to improve the identification of strong lensing systems.  The
corresponding AUC is listed in the legend.

% Model performance summary
The model performance for the mock HST data is AUC=0.975 for images
with the lens galaxy, and AUC=0.98 for images without the lens galaxy
(red solid and dashed).  On the other hand, the model for our
LSST-like dataset for one year has an AUC=0.625 with the lens galaxy
and AUC=0.579 without the lens galaxy (blue solid and dashed), and the
model for our LSST-like dataset for 10 years has an AUC=0.809 with the
lens galaxy and AUC=0.792 without the lens galaxy (green solid and
dashed).  Removal of the lens galaxy does not systematically perform
better, and is actually dependent on the size of the training set.  We
discuss relative model performance and complexity for images with and
without the lens galaxy in Section~\ref{sec:trainsetsize}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/ROC/ROC_compilation.pdf}
\caption{Red, blue, green: ROC curves for models trained on our whole
  10,000 training set and tested on our holdout set of 1,000.  These
  respectively correspond to the HST, LSST 1 year, and LSST 10 year
  data.  The solid lines are for data that include the lensing central
  galaxy, and the dashed lines for the data where there is no lensing
  galaxy, mimicking an ideal removal of the lens.  Model performance
  can be summarized by the area under the curve (AUC), labeled in the
  legend. $AUC=1$ is a perfect model, and $AUC=0.5$ is a useless
  model.}\label{fig:ROCcompilation}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/PR/PR_compilation.pdf}
\caption{Red, blue, green: Precision-recall (PR) curves for models
  trained on our whole 10,000 training set and tested on our holdout
  set of 1,000.  These respectively correspond to the same models and
  data shown in Figure~\ref{fig:ROCcompilation}.  An ideal model would
  reach both a precision (purity) and recall (completeness) that equal
  1.  Note, that this performance describes a data set with a 50-50
  split between lens and non-lens containing
  images.}\label{fig:PRcompilation}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\subsection{Effects of Regularization on Model Performance}\label{sec:regularization}
%-------------------------------------------------%

% Logistic regression and regularization
As described in Section~\ref{sec:LR}, LR trains a model with
complexity determined by the regularization parameter coefficient,
$C_\text{LogReg}$.  Larger values of $C_\text{LogReg}$ are less
regularized and allow for increased model complexity.  The highest
values of $C_\text{LogReg}$ will better describe features in the
training set.  However, an overly complex model will overfit the
training set at the expense of its performance on any independent test
set.  The regularization parameter ultimately defines the model
performance, and we must perform a parameter search to identify the
optimal value for $C_\text{LogReg}$.

%  Our running over C - score and AUC as characterizations
Figure~\ref{fig:regularization} shows the model performance as a
function of regularization parameter for each data set HST, \edit{LSST-best}, and
LSST10.  The solid and dotted blue lines respectively correspond to
the model performance on the test and training set, with the AUC as a
metric for performance.  In red, we show the train time as a function
of $C_{LogReg}$.

% Training behavior with C
As expected, the training set AUC increases and asymptotes with
$C_\text{LogReg}$.  With increasing model complexity, the model better
fits the training data set.  This is analogous to fitting a seventh
order polynomial to seven data points, where the fitting function will
go through every point but will not likely predict additional points.
With increasing model complexity, we are able to better able to
capture features that are generally characteristic of strong lensing
systems with arcs. However, past a certain $C_\text{LogReg}$, the
model performance on the test set decreases or asymptotes, as it has
overfit the training set.  We use the scaling of AUC with
$C_\text{LogReg}$ when training 8,000 out of our full 10,000 training
data set to determine the best value for $C_\text{LogReg}$.  However,
the optimal parameter is also dependent on the size of the training
set (see Section~\ref{sec:trainsetsize}), so this choice is not
generalizable.

% Train time with Creg
For fixed training set size, the log of the train time roughly scales
linearly with the log of $C_\text{LogReg}$.  Since lower values of
$C_\text{LogReg}$ correspond to a more regularized model, there is a
smaller volume in hyperparameter space to search for the best fit
coefficients.  The solution, on average, will converge more quickly,
for more regularized models.  The scaling is not purely monotonic
because the fitting still has some randomness associated with the path
it takes to convergence.

% Scores vs. regularization on larger train/test set with comparable
% grid search in parameterization
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
\begin{center}
\includegraphics[width=.66\columnwidth]{figures/Creg/HST.pdf}
\includegraphics[width=.66\columnwidth]{figures/Creg/LSST-best.pdf}
\includegraphics[width=.66\columnwidth]{figures/Creg/LSST10.pdf}
\caption{AUC of the model with varying LR regularization coefficient
  parameter, $C_{LogReg}$, used when training the model classifier.  We
  use a subset of the 10,000 training images to search over the LR
  $C_{LogReg}$ parameter, training on 8,000 and testing on 1,000.  Each
  panel corresponds to a different mock observation.  From left to
  right: HST, LSST for one year, and LSST for 10 years.  The solid
  blue lines correspond to the AUC of the test set, and the dotted
  blue lines to the AUC of the training set.  To avoid overfitting, we
  choose the smallest parameter for which the AUC of the test set is
  maximal: 5000, 10, and 5000, respectively.  In thin red solid lines,
  we show the train time of the model, which roughly increases in a
  log-log scaling with logistic regression coefficient
  parameter.}\label{fig:regularization}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\subsection{Dataset Size Dependence}\label{sec:datasize}
%-------------------------------------------------% 
%-------------------------------------------------% 
\subsubsection{Effects of Training Set Size on Performance and Train Time}\label{sec:trainsetsize}
%-------------------------------------------------%

In this section, we show the effects of training set size on model
performance on the hold-out test set of 1,000 images.  We also compare
the improvement between images that include the lens galaxy and images
with no lens galaxy.

% LSST10 and nLSST10 behavior
Figure~\ref{fig:trainsizeLSST} shows how the AUC depends on the log of
the size of the training set for both the LSST10 (\edit{LSST-best}) data in the
solid blue (red) line, and the nLSST10 (n\edit{LSST-best}) data in dashed blue
(red) line.  The AUC for models trained on the LSST10 data improves
almost linearly with the log of the training set size, increasing from
AUC=0.705 to AUC=0.788 when the train size is increased from
$2\times500$ lens/non-lens images to $2\times8000$ lens/non-lens
images.  However, for the nLSST10 data, where the lens galaxy has been
removed from the images, the improvement is less dramatic.  With the
same increase in training size, the AUC for nLSST10 changes from just
below 0.77 to just below 0.78.

% Comparison of the two LSST10 vs. nLSST10
In the LSST10 case, the trained model can incorporate the additional
information of the edges from the lens galaxy, which is correlated
with the lensing cross-section and likelihood of the image being a
lensing system.  The nLSST10 images do not contain this information,
but provide cleaner signals of the lensed image for lensed images that
occur close to the lens galaxy.  The cleaner signal in nLSST10 allows
for better model performance for smaller training data set sizes
($N_\text{train}\sim5,000$).  However, models trained on LSST10
improve more rapidly with train size, since the additional information
from the lens galaxy better describes the lens containing images.  For
$N_\text{train}\gsim 5,000$, the models trained on LSST10 outperform
those trained on nLSST10.

% LSST1 and nLSST1 behavior
In red, we see an analogous improvement for \edit{LSST-best} and n\edit{LSST-best}.
Here, the crossover happens at train size of $2x1000$, where an
increase in larger training data set size does not help improve the
AUC for n\edit{LSST-best}.

%M Model complexity, train time, and performance
Since the models that contain information from the lens galaxy edges
in LSST10 are more complex, the models require a larger training set
size for a better fit.  While model performance for LSST10 appears to
steadily increase, this comes at the cost of increased train time,
which is two-fold.  The train time will increase due to both an
increase in data to fit, and also an increase in optimal
$C_\text{LogReg}$ where the volume of hyperparameter space for allowed
solutions is larger (see red lines in
Figure~\ref{fig:regularization}).

% Train time dependence
We illustrate the dependence of train time on both the size of the
training set and model complexity in Figures~\ref{fig:traintimeLSST10}
and \ref{fig:traintimeLSST1}.  In red,
Figure~\ref{fig:traintimeLSST10} shows that the optimal values for
$C_\text{LogReg}$.  For LSST10, $C_\text{LogReg}$ roughly scales
logarithmically with the log of the train size, with an exception of
the data point corresponding to train size of 2000.  Generally, a
larger training set allows for an increase in model complexity without
reducing its ability to generalize.  This is also true for the optimal
$C_\text{LogReg}$ dependence on the number of training images in the
nLSST10 data.  But, the required complexity is systematically less
than for the LSST10 images.

% Train time
In blue, Figure~\ref{fig:traintimeLSST10} also shows the train time
for LSST10 and nLSST10 as a function of the size of the training set
for the optimal regularization parameter for that subset of the
training data.  Each model uses features extracted with the same HOG
parameterization from the grid search and the optimal regularization
parameter for that subset of the training data.  The train time of a
given model generally increases for increasing regularization
parameter.  For the subset of train size $N_\text{train}=2000$ in
LSST10, the optimal regularization parameter happened to be
$C_{LogReg}=100$, whereas it was $C_{LogReg}=500$ for the subset of
trainsize 1000, and $C_{LogReg}=1000$ for the subset of trainsize
4000.  This makes the train time increase at train size
$N_\text{train}=2000$ for LSST10 less dramatic than the average
log-log slope of approximately 2.

% Train time behavior for nLSST10
Since nLSST10 does not contain the lens galaxies, fewer of the
extracted features describe the lens system, requiring decreased model
complexity.  The increase in train time for nLSST10 as a function of
train size is mostly due to only having more data to fit in the
regression, leading to a steady and slow increase of train time with
number of training images with log-log slope of approximately 1.

% Train time behavior for LSST1 and nLSST1
For \edit{LSST-best}, Figure~\ref{fig:traintimeLSST1} shows the same
relationships between train time and train size in blue, and the
optimal regulariztion parameter (or model complexity) in red.
Contrary to what we found when comparing LSST10 to nLSST10, n\edit{LSST-best}
requires more model complexity than \edit{LSST-best}.  Recall that \edit{LSST-best} and
n\edit{LSST-best} correspond to single epoch simulated exposures with the best
seeing; these images exhibit better resolution but worse signal to
noise.  The switch in required model complexity corresponds to a
trade-off between features from the lens galaxy providing additional
information or swamping the signal from a strong lensing arc.

% AUC vs. regularization on larger train/test set with comparable
% grid search in parameterization
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/trainsize/combined_LSST10_with_nLSST10.pdf}
\caption{Solid (dotted) blue line: AUC for models with varying size of
  the training set for LSST10 (nLSST10).  Solid (dotted) red line:
  Same for \edit{LSST-best}. The improvement of AUC scales roughly linearly with
  the log of the training set size.  However, LSST10 and \edit{LSST-best} have a
  steeper improvement with data training size.  The performance of
  nLSST10 and n\edit{LSST-best} models trained on smaller training data sets is
  better than the respective LSST10 and \edit{LSST-best} models trained on the
  same size data, but the LSST10 and \edit{LSST-best} models outperform with
  larger size training data set.}\label{fig:trainsizeLSST}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]
\begin{center}
  \includegraphics[width=\columnwidth]{figures/traintime/LSST10.pdf}}
\caption{Solid (dotted) blue line: Train time for models with varying
  size of the training set for LSST10 (nLSST10).  Train time roughly
  scales logarithmically with the train size, but the train time is
  also affected by model complexity.  Solid (dotted) red line: Best
  regularization parameter as a function of train size.  Note, LSST10
  requires more model complexity to exceed the performance of nLSST10
  (see blue solid and dotted lines in
  Figure~\ref{fig:trainsizeLSST}), and therefore requires more
  training time for continual increase in
  performance. }\label{fig:traintimeLSST10}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]
\begin{center}
  \includegraphics[width=\columnwidth]{figures/traintime/LSST-best.pdf}}
\caption{Same as Figure~\ref{fig:traintimeLSST10}, but for \edit{LSST-best}.
  Here, n\edit{LSST-best} requires more model complexity than \edit{LSST-best}.  Note, that
  \edit{LSST-best} and n\edit{LSST-best} have better seeing but reduced overall signal to
  noise than LSST10 and nLSST10.  This corresponds to better
  resolution, and therefore sharper edge features that prominently
  corespond to arcs.}\label{fig:traintimeLSST1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------% 
\subsubsection{Effects of Rotation on AUC}
%-------------------------------------------------%

% Augmenting dataset
To augment our training set, we rotated each image in the set by
multiples of $90^o$.  Since our feature extraction method of HOG is
not rotationally invariant, augmenting our data by a factor of four
naturally optimizes the use of available training data.  This has an
equivalent improvement to the study illustrated in
Figure~\ref{fig:trainsizeLSST}.

% Rotations of the test set
We also tested the effects of evaluating our model on all four
rotations of the test set, and using the average score of each test
image to calculate the AUC.  In Figure~\ref{fig:rotation_test}, we
show the AUC for different orientations of the datasets.  The x-axis
corresponds to each of our three datasets, with and without the lens
galaxy.  The y-axis shows the AUC.  The filled blue circles correspond
to the four AUCs calculated when the model evaluated images at each of
the four rotations.  The filled red stars correspond to the AUC
calculated from the average of all four test scores, which are
systematically higher than any one rotation.  The average score across
all rotations for each image is likely to be less noisy for the whole
test sample, giving an improved AUC.

% Summary figure
Figure~\ref{fig:rotation_test} also summarizes the best-case results
of our models trained on our entire 10,000 training sets, and tested
on our holdout 1,000 test set.  Recall, however, that we expect model
performance on images containing the lens galaxy to improve further
with larger training sets (see Figure~\ref{fig:trainsizeLSST}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/rotated_aucs/rotated_aucs.pdf}
\caption{Our summary figure: The AUCs of models trained on the full
  10,000 and tested on the holdout 1,000.  Blue circles: AUC
  calculated from scores of images at a given rotation (e.g. 0, 90,
  180, and 270 degrees).  Red stars: AUC calculated from the average
  score of all rotations of each image.  The average score produces an
  improved AUC in all data sets.  We expect the AUC to further improve
  with increased train size.}\label{fig:rotation_test}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\subsection{Image Classification Performance}\label{sec:performance}
%-------------------------------------------------%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
\begin{center}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/tp.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/fp.pdf}\\\vspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/bp.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/bn.pdf}\\\vspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/fn.pdf}\hspace{5pt}
\includegraphics[width=.8\columnwidth]{figures/tiles/compilations/LSST10/tn.pdf}\\
\caption{LSST 10 year mock images. Left two columns: Lens containing
  images, annotated with the image score assigned by our trained
  classifier.  Right two columns: Non-lens containing images,
  annotated with the image score.  The top two rows show
  characteristic images that will be accepted with a high threshold
  for classification, contributing to the bottom left of the ROC curve
  in Figure~\ref{fig:ROCcompilation}.  The middle two rows show
  characteristic images that will be accepted with a moderate
  threshold, contributing to the knee of the ROC curve, with true
  positive rates and false positive rates of $tpr\approx0.8$ and
  $fpr\approx0.25$, respectively.  The bottom two rows show
  characteristic images that will only be accepted with an extremely
  lenient threshold, contributing to the top right area of the ROC
  curve.}\label{fig:ROCsamplesLSST10}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Populating the ROC Curve}
% Defining the 6 paradigms of classification
In this section, we discuss the different image types that our model
is most and least able to successfully classify.  We have six
paradigms of model performance on the mock images based on the score
an image receives when evaluated by the trained model, and its true
label.  From highest scoring to lowest scoring: True Positives (tp),
False Positives (fp), Borderline Positives (bp), Borderline Negatives
(bn), False Negatives (fn), and True Negatives (tn).

Figure~\ref{fig:ROCsamplesLSST10} illustrates four images from each
paradigm for the hold-out test set of LSST10.  The trained model used
the entire 10,000 image training set.  The left two columns show lens
systems, and the right two columns show non-lens systems.  The
annotation in the top left corner of each images shows the score.

% tp, tn description
In general, the true positives (the highest scoring lens systems) have
lensed images with large magnification.  The the true negatives (the
lowest scoring non-lens systems) have small lens galaxies with
galaxies along the line of sight that are rounded.  The successful
classification of these two paradigms are least sensitive to the
threshold.  On the other hand the failed classification of the false
positives (the highest scoring non-lens systems) and the false
negatives (the lowest scoring lens systems) are also least sensitive
to the threshold.  False negatives are typically lens systems with
lensed images of smaller magnification and minor distortions that
mimic along the line of sight galaxies that the model has learned to
ignore.  False positives are often non-lens systems with elongated,
elliptical, or ``fuzzy'', galaxies along the line of sight whose signal
blends with the lens galaxy contributing to the false positive rate
even for conservative thresholds.  Visually, these false positives are
virtually indistinguishable from true arcs, and would require
spectroscopic follow-up.

% Borderlines
The middle two rows of Figure~\ref{fig:ROCsamplesLSST10} illustrate
the borderline positives and borderline negatives. The successful
classification of the borderline positives and negatives are most
sensitive to the threshold, and would be the first candidates for
alternative classification methods, such as visual follow-up.
Thresholds set around these scores yield a true positive rate and
false positive rate of $tpr\approx0.8$ and $fpr\approx0.25$,
respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
\subsubsection{Dependence on Lens-Model Parameters}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Summary of subsection
Here, we examine lens-model parameters that affect how well our
pipeline can classify the system.  The lens-model parameters we
examined are the redshift, ellipticity, orientation angle, and
velocity dispersion of the lensing galaxy, and also the magnification
of the lensed image compared with its original size.  We found that
the magnification of the lensed image is the most correlated
lens-model parameter with our trained model performance, with a
secondary and related correlation with lens galaxy velocity dispersion
that is encapsulated in the Einstein radius.  The more strongly lensed
an image is, the larger its magnification, and the easier it is for
our trained model to classify.

% Discuss figure
In Figure~\ref{fig:scorevslensparams}, we show the classification
\edit{score as a function of Einstein radius and image magnification
  of the source galaxy in each of our samples.} In either an HST-like
or LSST-like observation, lensing systems with images that have
magnification $\gsim7$ will almost certainly be classified as positive
with threshold scores above $\gsim0.5$.  These systems are also those
that are most easily classified by eye.  However, our trained model
has varying performance for systems with lower magnifications.

% Some dependence on redshift and velocity dispersion, which is folded
% into magnitude
In our mock HST sample, all but one of our lensing systems with
magnifications $\lsim7$ will be successfully identified for
classification threshold scores $\gsim0.5$.  These highly magnified
systems typically have lenses that are at higher redshifts of and/or
lens galaxies with velocity dispersions larger than
$\sigma_v\gsim230.$~km/s.  Note that lensing galaxies with smaller
velocity dispersions are less massive and therefore have a smaller
efficiency of lensing cross section.  The smaller efficiency means
that a background galaxy is less likely to be strongly lensed with
high magnification.  Also, due to hierarchical structure formation,
galaxies are less massive at higher redshifts, so the trends of model
performance with these three parameters are somewhat degenerate with
one another.

The relationship between our model performance and these lens
parameters indicates that the magnification of lensed images is the
most relevant, and the distribution of image magnification in a data
set will impact trained model performance.  We did not find strong
correlations in model performance with other lens parameters.  It is
also useful to keep in mind that lensed galaxy images with
magnification $\lsim 5$ are often visually indistinguishable from
\edit{edge-on disk galaxies along the line of sight, which can lead to
  false positives, features from which our model has learned to
  downweight.  Note that the model would be more sensitive to systems
  with lower magnification without galaxies along the line of sight.}

In a forthcoming paper, we will discuss how class imbalance, or
differences between the lens-model parameter distributions in the
training and test sets affect model performance and will explore a
method to correct for this.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]
  \centering 
  \mbox{
    \subfigure{\includegraphics[width=.66\columnwidth,trim=13 0 10 0,
        clip]{figures/lens_parameter_dependence/HST_Einstein_Radius_z.pdf}}\hfill
    \subfigure{\includegraphics[width=.66\columnwidth,trim=13 0 10 0,
        clip]{figures/lens_parameter_dependence/LSST10_Einstein_Radius_z.pdf}}\hfill
    \subfigure{\includegraphics[width=.66\columnwidth,trim=13 0 10 0,
        clip]{figures/lens_parameter_dependence/LSST-best_Einstein_Radius_z.pdf}\hfill}
  }\\ 
  \mbox{
    \subfigure{\includegraphics[width=.66\columnwidth,trim=13 0 10 0,
        clip]{figures/lens_parameter_dependence/HST_Magnification_z.pdf}}\hfill
    \subfigure{\includegraphics[width=.66\columnwidth,trim=13 0 10 0,
        clip]{figures/lens_parameter_dependence/LSST10_Magnification_z.pdf}}\hfill
    \subfigure{\includegraphics[width=.66\columnwidth,trim=13 0 10 0,
        clip]{figures/lens_parameter_dependence/LSST-best_Magnification_z.pdf}\hfill}
  }\\ 
\caption{Left to right: Image classification score on HST, LSST10,
  \edit{LSST-best} mock observed lensing images as a function of Einstein
  radius (top) and magnification (bottom). Magnification is the
  strongest indicator of how likely a lensing system will be
  successfully classified.  The Einstein radius has a weaker
  correlation with how easily a lensing system might be classified,
  since it contains information on both the velocity dispersion of the
  lens and the redshifts of the lens and source galaxies.  High
  velocity dispersion lens galaxies are likely to produce high
  magnification images of the source galaxies.  }\vspace{10pt}
\label{fig:scorevslensparams}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------%
\section{Summary and Discussions}
\label{sec:conclusions}
%-------------------------------------------------%

% Summarize paper
We have presented a supervised classification pipeline to
automatically identify galaxy-galaxy strong lensing systems using a
histogram of oriented gradients (HOG) as a feature extractor, and
logistic regression (LR) as a machine learning algorithm.  Our
pipeline can easily be extended to identify other strong lensing
features, such as multiply imaged quasars, and to test alternative
features and/or machine learning algorithms.

% Training set
We have also made use of a new sophisticated set of mock observations,
which will be made publicly available.  The lensing systems have
lens galaxies generated with a realistic redshift distribution and
along the line of sight galaxies drawn from Hubble Ultra-Deep field
observations.  We have explored results from mock HST, one year LSST,
and ten year LSST observations.

We summarize key points below:

%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}

\item We have designed our pipeline to easily select and add image
  pre-processing and feature extraction methods, and to select a
  machine learning algorithm for classification.  Additionally, the
  user can easily perform parameter searches to train a model with the
  best parameters for a given problem.

\item We have tested and run parameter tests for a histogram of
  oriented gradients as an efficient and effective feature extractor
  for galaxy-galaxy strong lensing systems in both a space based
  (HST-like) and ground based (LSST-like) observation.  We have also
  tested and run parameter tests for logistic regression as a
  scalable, cheap, and effective machine learning algorithm.

\item We find AUC values of ROC curves of optimized classifier models
  to yield $AUC=0.975$ for the HST-like data, $AUC=0.809$ for the
  stacked LSST-like data.  Model performance exhibits continual
  increase with the training size.

\item While removal of the lens galaxy improves model performance for
  smaller size training samples, features from the lens galaxy improve
  model performance for larger training data sets.

\item Images that were easiest for our model to classify typically
  were lens systems that had high lensed image magnification and a
  lens galaxy with large velocity dispersion or non-lens systems with
  lens galaxies with smaller velocity dispersion and non-elongated
  along the line of sight galaxies.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%

% Benefits of our method
We emphasize that simple linear classifiers, such as logistic
regression, are scalable and relatively easy to parallelize with open
source tools such as {\em Apache Spark}\footnote{http://spark.apache.org/}.
Our work indicates that HOG feature extraction plus a linear
classifier captures much of the complexity in the arc finding problem.
{\it These results indicate that effective image classification of
  galaxy-galaxy strong lensing systems can be achieved without large
  scale computing resources, or specialized hardware like GPUs.}
The methods also scale to large datasets on a computing cluster, if
needed.

% Caveat of the mock data set
One major caveat to our resuls is the fact that our mock data likely
does not describe the full range of lens and non-lens images that will
be observed.  For example, the along the line of sight galaxies in our
training and test images all come from the Hubble Ultra Deep field,
sampling a smaller range of potential contaminants that are not
associated with a lensed image.  Also, the redshift of the source
galaxies have been fixed to $z_s=2$.  We will address these
shortcomings in subsequent work.

% Interpreting TP and FP rates
Finally, we must take the class balance, or relative number of lens to
non-lens systems, into account when assessing a metric.  The
Purity-Recall (purity-completeness) metric is sensitive to the ratio
of lens and non-lens systems in the data.  Both of our training and
test data sets have $50\%$ lensed and $50\%$ unlensed images, which is
not expected in observations.

% Realistic ratios
\edit{For images selected for a massive elliptical, the number of non-lenses
will outnumber lenses by at least 1000 to 1.  Therefore, a sample that
is $50\%$ pure requires a classifier with a false positive rate of
0.001.  Looking at the green line in Figure~\ref{fig:PRcompilation}, a
classification threshold for a pure sample only encapsulates
$\sim10\%$ (in recall) of all of our lenses before contamination from
non-lenses leaks into the selection.  While information from other
bands will certainly improve the model performance, a maximally large
and pure sample from our method would likely require further
filtering, e.g. by citizen science.}

% ROC curve metric
The ROC curve metric is insensitive to the ratio, but is sensitive to
the sampling.  Given alternative lens and non-lens sample splittings,
our true positive and false positive rates in the ROC curves would
stay the same, making the ROC curve a more standard metric in the
literature.  On the other hand, the ROC curves show a representative
rate for lens and source distributions that are evenly sampled.  We do
{\it not} expect this sampling to be representative of what we might
expect from a an observational survey.  We leave these additional
challenges to future work.

\acknowledgments CA acknowledges support from the Enrico Fermi
Institute at the University of Chicago, and the University of Chicago
Provost's Office. NL would like to thank the funding support from
NSFC, grant no.11503064, and Shanghai Natural Science Foundation,
grant no. 15ZR1446700. This work was also supported in part by the
Kavli Institute for Cosmological Physics at the University of Chicago
through grant NSF PHY-1125897 and an endowment from the Kavli
Foundation and its founder Fred Kavli.
\lastpagefootnotes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%\section{Sample stacked mock LSST images}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%\subsection{BCG included}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{ms}

\end{document}